% mainfile: ../../../../master.tex
\subsection{AOSP}
\label{task:20231115_aosp}

\subsubsection*{Starting Android System Server and Other Apps using Zygote}

During its initialization, Zygote will check for \texttt{start-system-server} flag, and if set, will bring up \texttt{SystemServer} in the following sequence:

\begin{longtable}{p{.25\linewidth}  p{.4\linewidth} p{.3\linewidth}} 
\toprule
Method 	& Description & Source\\
\midrule
\endhead

\url{ZygoteInit.forkSystemServer} 
& Runs after the Zygote process has been initialized. It is hardcoded with System Server classpath\footnote{\url{com.android.server.SystemServer}, the source for which is stored in \url{AOSP/frameworks/base/services/java/com/android/server/SystemServer. java}.} as one of the arguments to call \texttt{Zygote.forkSystemServer} to spawn \texttt{SystemServer} process.
&\path{AOSP/framework/base/core/java/com/android/internal/os/ZygoteInit.java}\\

\url{Zygote.forkSystemServer} 
& \texttt{Zygote} class wraps native methods that communicate with Android Runtime, one of whom is \texttt{com\_android\_internal\_os\_nativeForkSystemServer}.
&\path{AOSP/framework/base/core/java/com/android/internal/os/Zygote.java}\\

\texttt{com\_android\_internal\_os\_nativeForkSystemServer}
& Calls \texttt{zygote::ForkCommon} and \texttt{SpecializeCommon} which does the actual forking.
&\path{AOSP/frameworks/base/core/jni/com_android_internal_os_Zygote.cpp}\\

\texttt{SpecializeCommon}
& Looks at the flags and Process ID for setting up sandboxing, configuring the correct SE Linux context, and process capabilities. Afterwards, it will call \texttt{Zygote} methods for post-fork procedures.
&\path{AOSP/frameworks/base/core/jni/com_android_internal_os_Zygote.cpp}\\

\url{Zygote.callPostForkSystemServerHooks}
& Calls \texttt{ZygoteHooks.java} at the end of specialization procedures. Only applicable for \texttt{SystemSever}.
&\path{AOSP/framework/base/core/java/com/android/internal/os/Zygote.java}\\

\url{Zygote.callPostForkChildHooks}
& Calls \texttt{ZygoteHooks.java} at the end of specialization procedures. Applicable to all applications and services including \texttt{SystemServer}
&\path{AOSP/framework/base/core/java/com/android/internal/os/Zygote.java}\\

\url{ZygoteHooks.postForkSystemServer} and \url{ZygoteHooks.postForkSystemServer}
& Wrappers for \texttt{ZygoteHooks} inside the Android Runtime. They call their respective native code inside the ART via Java Native Interface.
&\path{AOSP/libcore/dalvik/src/main/java/dalvik/system/ZygoteHooks.java}\\

\url{ZygoteHooks_nativePostForkSystemServer}
& Loads the specialized class libraries to start the the System Server. 
&\path{AOSP/art/runtime/native/dalvik_system_ZygoteHooks.cc}\\

\url{ZygoteHooks_nativePostForkChild}
& Loads the specialized class libraries to start the service/application. 
&\path{AOSP/art/runtime/native/dalvik_system_ZygoteHooks.cc}\\

\url{handleSystemServerProcess}
& The control returns to \texttt{ZygoteInit}, and finish remaining work for the newly forked system server process, and calls \texttt{ZygoteInit.zygoteInit}.
&\path{AOSP/framework/base/core/java/com/android/internal/os/ZygoteInit.java}\\

\url{ZygoteInit.zygoteInit}
& The main function called when started through the zygote process, which calls \texttt{RuntimeInit.applicationInit} 
&\path{AOSP/framework/base/core/java/com/android/internal/os/ZygoteInit.java}\\

\url{RuntimeInit.applicationInit}
& Calls the \texttt{public static void main} method of the application
% \footnote{Here all the Android services are started: \texttt{PackageManagerService}, \texttt{ActivityManagerService}, and so on.}. 
&\path{AOSP/framework/base/core/java/com/android/internal/os/RuntimeInit.java}\\

\url{ZygoteServer.runSelectLoop}
& After forking has finished, the control enters \texttt{ZygoteServer} which starts an endless loop that handles incoming connections with \texttt{ZygoteConnection.processCommand}.
&\path{AOSP/framework/base/core/java/com/android/internal/os/ZygoteServer.java}\\

\url{ZygoteConnection.processCommand}
& Calls \texttt{Zygote.forkAndSpecialize}
% \footnote{Old code to be replaced soon.} 
which is a version of \texttt{ZygoteInit.forkSystemServer} for the masses.
&\path{AOSP/framework/base/core/java/com/android/internal/os/ZygoteConnection.java}\\

\url{Zygote.forkAndSpecialize}
& A version of \url{Zygote.forkSystemServer} for the masses. 
&\path{AOSP/frameworks/base/core/jni/com_android_internal_os_Zygote.cpp}\\

\texttt{com\_android\_internal\_os\_nativeForkAndSpecialize}
& Calls \texttt{zygote::ForkCommon} and \texttt{SpecializeCommon} which does the actual forking, and returns to \texttt{ZygoteInit} and immediately enters \texttt{ZygoteServer}.
&\path{AOSP/frameworks/base/core/jni/com_android_internal_os_Zygote.cpp}\\

\midrule
\caption{System Server and Applications Initialization Sequence} 
\label{tab:sysserandappinitializationsequence}
\end{longtable}